---
title: "Brugmansia: Learning a nonzero amount about ecological data"
date: "2018-03-21"
slug: "Brugmansia-Learning-a-Nonzero-Amount-About-Ecological-Data"
draft: true
---


```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      echo = TRUE, 
                      dpi = 180)
library(BIEN)
library(dplyr)
library(ape)
library(maps)
library(sp)
library(stringr)
```


## Introduction
This post is inspired by the imbalance of my interest in the field of ecology and my understanding of it. Because I have some knowledge of statistics and programming (the work I do is occasionally referred to as "data science" by people less resentful than me), I thought I would do some simplistic data stuff! I'm going to assume anyone reading this has as much knowledge of ecology as I do (i.e. they read a book about it years ago that they liked very much).

## Brugmansia
I chose to focus on the plant genus (the step above species in taxonomic rank) [Brugmansia](https://en.wikipedia.org/wiki/Brugmansia) for my exploration. It has a really interesting and complicated history worth checking out!

## Obtaining Data
**Insert information about BIEN**
  
``` {r import-data}
brugmansia_data <- 
  BIEN_occurrence_genus(genus = "Brugmansia",
                        cultivated = TRUE,
                        only.new.world = FALSE,
                        all.taxonomy = TRUE,
                        native.status = TRUE,
                        observation.type = TRUE,
                        political.boundaries = TRUE) %>% 
  filter(!is.na(latitude),
         !is.na(longitude),
         !is.na(date_collected))

brugmansia_data_col <-
  brugmansia_data %>% 
  filter(country == "Colombia")

map("world", fill = TRUE, col = "grey", bg = "light blue")
points(cbind(brugmansia_data$longitude,
             brugmansia_data$latitude),
       col = "red", pch = 20, cex = 1)

```
