---
title: "Brugmansia: Learning a nonzero amount about ecological data"
date: "2018-03-21"
slug: "Brugmansia-Learning-a-Nonzero-Amount-About-Ecological-Data"
draft: true
---


```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE, 
                      echo = TRUE, 
                      dpi = 180)
library(BIEN)
library(dplyr)
library(ape)
library(maps)
library(sp)
library(leaflet)
library(stringr)
```


## Introduction
This post is inspired by the imbalance of my interest in the field of ecology and my understanding of it. So I thought I would do some simplistic data stuff as an initial introduction into the world. I'm going to assume anyone reading this has as much knowledge of ecology as I do (i.e. they read a book about it years ago that they liked very much).

I chose to focus my exploration on the plant genus (the step above species in taxonomic rank) [Brugmansia](https://en.wikipedia.org/wiki/Brugmansia), specifically in the country of Colombia. Brugmansia has a really interesting and complicated history worth digging into.

## Obtaining Data
**Insert information about BIEN**
  
``` {r import-data}
brugmansia_data <- 
  BIEN_occurrence_genus(genus = "Brugmansia",
                        cultivated = TRUE,
                        only.new.world = FALSE,
                        all.taxonomy = TRUE,
                        native.status = TRUE,
                        observation.type = TRUE,
                        political.boundaries = TRUE) %>% 
  filter(!is.na(latitude),
         !is.na(longitude),
         !is.na(date_collected))

brugmansia_data_col <-
  brugmansia_data %>% 
  filter(country == "Colombia")

leaflet(data = brugmansia_data_col,
        width = 750,
        height = 500) %>% 
  addProviderTiles(providers$OpenTopoMap) %>% 
  addCircleMarkers(~longitude, ~latitude,
                   clusterOptions = markerClusterOptions(),
                   popup = ~paste0("Latitude: ", as.character(latitude), "</br>",
                                   "Longitude: ", as.character(longitude), "</br>",
                                   "Locality: ", as.character(locality)),
                   stroke = FALSE,
                   fillOpacity = 0.66)

```
